A sheaf--function dictionary for algebraic tori over local fields

### 2013-10-10 shanghai.tex
Adding stuff about localizations and local systems to section 6

A Weil sheaf $(\gqcs{L},\phi)$ on $G$
may be interpreted as a constructible $\ell$-adic sheaf $\gqcs{L}$ on $\bG$ together with
an action of the Weil group $\Weil{}$ on $\gqcs{L}$ compatible with the action of
$\Gal(\bFq/\Fq)$ on $\bG$. We now briefly review this point of view for later use in
Section~\ref{sec:basechange}; see also \cite{deligne:80a}*{Def.~1.1.10}.  If $w \ceq \Frob{}^n$ we write $w_G$ for $\Frob{G}^n$.
For each such $w$ we define an isomorphism $\varphi(w) : w_G^* \gqcs{L}\to \gqcs{L}$ by
$
\varphi(w) \ceq  \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi)
$; 
these isomorphisms satisfy $\varphi(uv) = \varphi(v) \circ v_G^* \varphi(u)$ for $u,v\in \Weil{}$, and
we have $\phi = \varphi(\Frob{})$. 
If we further define $\varphi_1(w) \ceq (w_G)_*(\varphi(w)^{-1})$ then
$\varphi_1(w) : (w_G)_* \gqcs{L}\to \gqcs{L}$ is an isomorphism and
$\varphi_1(uv) = \varphi_1(u) \circ (u_G)_* \varphi_1(v)$ for $u,v\in \Weil{}$.
Thus the pair $(\gqcs{L},\varphi_1)$ almost satisfies the criteria for an action of $\Weil{}$
on $\gqcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ in the sense of
\cite{deligne:SGA4.5}*{Expos\'e XIII, 1.1}, failing only because $\Weil{}$ is not profinite.
\Clifton{Improve this.}
As we will see, our use of $\Weil{}$ rather than $\Gal(\bFq/\Fq)$ allows quasicharacter sheaves to
correspond to unbounded characters.

@@

Now we localize $C(G)$.
Let us say that $(g,T)$ in $C(G)$ is a \emph{weak isomorphism} if $g$ is surjective
and $T$ is an isomorphism of vector spaces.
\Clifton{Should this be another lemma in this section?}
The set $W$ of weak isomorphisms in $C(G)$
is a saturated, left multiplicative system in $C(G)$ in the sense of
  \cite{kashiwara-schapira:CatgoriesSheaves}*{Defs 7.1.5, 7.1.19}.
Let \cdef{$C(G)[W^{-1}]$} be the localization of $C(G)$ at $W$.
A morphism in $C(G)[W^{-1}]$ is an equivalence class of a diagram
%  \begin{equation}\label{eq:morGCf}
\[   
\begin{tikzcd}[column sep=30]
   (f,\psi) \arrow{r}{(g,T)} & (f',\psi') & \arrow[swap]{l}{(h,U)} (f_1,\psi_1)
   \end{tikzcd}
\]
%  \end{equation}
 where $(g,T) : (f,\psi) \to (f',\psi')$ is any map in $C(G)$ 
 and $(h,U) :  (f',\psi') \to (f_1,\psi_1)$ is a weak isomorphism in $C(G)$; see \cite{kashiwara-schapira:CatgoriesSheaves}*{Thm 7.1.16} for more detail, including the equivalence relation mentioned above.
%Likewise, $C(G)[W^{-1}]$ is a tensor category.

@@

>>>>

A Weil sheaf $(\gqcs{L},\phi)$ on $G$
may be interpreted as a constructible $\ell$-adic sheaf $\gqcs{L}$ on $\bG$ together with
an action of the Weil group $\Weil{}$ on $\gqcs{L}$ compatible with the action of
$\Gal(\bFq/\Fq)$ on $\bG$. We now briefly review this point of view for later use in
Section~\ref{sec:basechange}; see also \cite{deligne:80a}*{Def.~1.1.10}.  If $w \ceq \Frob{}^n$ we write $w_G$ for $\Frob{G}^n$.
For each such $w$ we define an isomorphism $\varphi(w) : w_G^* \gqcs{L}\to \gqcs{L}$ by
$
\varphi(w) \ceq  \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi)
$; 
these isomorphisms satisfy $\varphi(uv) = \varphi(v) \circ v_G^* \varphi(u)$ for $u,v\in \Weil{}$, and
we have $\phi = \varphi(\Frob{})$. 
If we further define $\varphi_1(w) \ceq (w_G)_*(\varphi(w)^{-1})$ then
$\varphi_1(w) : (w_G)_* \gqcs{L}\to \gqcs{L}$ is an isomorphism and
$\varphi_1(uv) = \varphi_1(u) \circ (u_G)_* \varphi_1(v)$ for $u,v\in \Weil{}$.
Thus the pair $(\gqcs{L},\varphi_1)$ almost satisfies the criteria for an action of $\Weil{}$
on $\gqcs{L}$ compatible with the action of $\Gal(\bFq/\Fq)$ in the sense of
\cite{deligne:SGA4.5}*{Expos\'e XIII, 1.1}, failing only because $\Weil{}$ is not profinite.
As we will see in Corollary~\ref{cor:bounded-and-finite}, our use of $\Weil{}$ rather than $\Gal(\bFq/\Fq)$
allows quasicharacter sheaves to correspond to unbounded characters.

@@

Now we localize $C(G)$.
Let us say that $(g,T)$ in $C(G)$ is a \emph{weak isomorphism} if $g$ is surjective
and $T$ is an isomorphism of vector spaces.  Recall from
\cite{kashiwara-schapira:CatgoriesSheaves}*{Def 7.1.5} that a \emph{left multiplicative system}
is a family of morphisms $W$ that
\begin{enumerate}
\item contains all isomorphisms,
\item is closed under composition,
\item is closed under pullback by arbitrary morphisms in $C(G)$,
\item and if the cokernel \David{Is $C(G)$ abelian?} of a morphism $\phi \in \mor C(G)$ is in
$W$ then so is the kernel of $\phi$.
\end{enumerate}

We further say that $W$ is \emph{saturated} if the following property holds for
morphisms $f, g$ and $h$ that can be composed appropriately:
whenever $g \circ f$ and $h \circ g$ are in $W$ then so is $h$
\cite{kashiwara-schapira:CatgoriesSheaves}*{Def 7.1.19}.

\begin{lemma}
The family of weak isomorphisms in $C(G)$ is a saturated, left multiplicative system.
\end{lemma}
\begin{proof}
Any isomorphism $(g, T)$ in $C(G)$ has both $g$ and $T$ isomorphisms, and thus is a weak isomorphism.
Clearly, the composition of two weak isomorphisms is a weak isomorphism.  The remaining properties
follow from the fact that in a weak isomorphism $(g, T)$, $g$ is an \'etale cover, together with standard
facts on \'etale covers and pullbacks. \David{Is this sufficient?}
\end{proof}
Let $W$ be the family of weak isomorphisms and \cdef{$C(G)[W^{-1}]$} be the localization of $C(G)$ at $W$.
A morphism in $C(G)[W^{-1}]$ is an equivalence class of diagrams
%  \begin{equation}\label{eq:morGCf}
\[   
\begin{tikzcd}[column sep=30]
   (f,\psi) \arrow{r}{(g,T)} & (f',\psi') & \arrow[swap]{l}{(h,U)} (f_1,\psi_1)
   \end{tikzcd}
\]
%  \end{equation}
 where $(g,T) : (f,\psi) \to (f',\psi')$ is any map in $C(G)$ 
 and $(h,U) :  (f',\psi') \to (f_1,\psi_1)$ is a weak isomorphism in $C(G)$ \cite{kashiwara-schapira:CatgoriesSheaves}*{Thm 7.1.16}.
%Likewise, $C(G)[W^{-1}]$ is a tensor category.

@@

To prove the next proposition, we will need a certain perspective on local systems.
 If $\qcs{E}$ is any local system on a connected scheme $G$ and $\bg$ is a geometric point on $G$,
 we may associate to $\qcs{E}$ a character of the \'etale fundamental group $\pi_1(G, \bg)$ \David{add citation}.
 Since $\pi_1(G, \bg)$ is defined as a projective limit of automorphism groups of finite \'etale covers, a general
 local system will correspond to a projective limit of characters of such groups.  But some local systems can be
 described in terms of a single cover $f: H \to G$ and a representation $\psi: \ker f \to \Aut(V)$;
 two such descriptions yield isomorphic local systems if and only if they are related by a weak
 isomorphism in $C(G)$.  While the description of local systems in terms of $\pi_1(G, \bg)$ does
 not generalize well to non-connected $G$, the description in terms of $C(G)$ does.  \David{Look okay?}
